Complete Algebraic File Synchronization

### 2015-12-22 2015/comparisons.tex
More improvements and references

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{centernot}
\usepackage{tikz}
\usepackage{cleveref}

\input{macros}

\title{Complete Algebraic File Synchronization}
\author{Elod Pal Csirmaz}

\begin{document}
\maketitle
\begin{abstract}
Abstract goes here
\end{abstract}

\input{intro}

\section{Definitions}\label{sec_def}

\input{def-filesystem}

\input{def-commands}

\input{axioms}

\section{Update Detection}\label{sec_update}

\input{upd-def}

\input{ordering}

\input{upd-correct}

\section{Reconciliation}\label{sec_rec}

\input{rec-intro}

\input{works-inverse}

\input{rec-sound}

\input{rec-complete}

@@

We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (see \cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

@@

% TODO Conclusions

\section{Further Research}

TODO: next steps. multiple replicas

\input{algebra}

\section{Acknowledgments}

I thank Laszlo Csirmaz for his invaluable input on previous drafts of this paper,
and Bill Zissimopoulos for his observations on algebraic file synchronization that triggered this research.

\begin{thebibliography}{99}

\input{references}

\end{thebibliography}

\end{document}

>>>>

@@

\subsection{Liberal and conservative Reconciliation}
We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (see \cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

@@

### 2015-12-22 2015/def-commands.tex
More improvements and references

Next we need to define commands on the filesystem.

... TODO ...

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(n,\valv)$, $edit(n,\valv)$ and $remove(n)$ 
where $n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
edits that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands.
Indeed, Bill Zissimopoulos suggested (TODO) that extending this distinction to more commands
ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
output \emph{value} of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

Notice, however, that the commands listed above also encode some information about 
their input, the state of the filesystem
before the command is applied. In particular, $create(n,\valv)$ requires that there are no files
or directories at $n$, while $edit(n,\valv)$ and $remove(n)$ require the opposite.
This creates an arbitrary asymmetry where
there is now more information encoded into commands about their output than about their input.
In line with the aim to encode as much information as possible in the commands, % TODO Where is this mentioned?
and in order to resolve this asymmetry, we propose a set of commands that encode
the type of the original state of $\FS(n)$ as well.
(Some real-life filesystem commands like $rmdir$ do this already.)

We therefore have four pieces of information related to each command:
its input type, its output type, the node to which it is applied, and finally
the value that it stores at the node.
We will denote commands using quadruples that list these,
and use $\cchard$, $\ccharf$ or $\ccharb$ to represent the different
input and output types. In this system, $\cbdaa{n}{\vald}$ 
is what represents $mkdir(n,\vald)$.
Alternatively, we use a triple of the two types and a node if 
there is only a single suitable output value for the given type (e.g. $\empt$).
For example, $\cdba{n}$ represents $rmdir(n)$.

@@

The exact behavior of our commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxynv$ where
$X$ and $Y$ are types in $\typeset$,
$n\in\setn$ and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
\[ \cxynv\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

@@

% Simplifications
% ---------------

>>>>

Next we define commands on the filesystem.
As described above, we aimed to select a set of commands
that capture as much information as possible
about the operations, and is also symmetric.
We start by summarizing our reasons for doing so.

Let us consider what kind of information is usually encoded in filesystem operations.
A usual set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$---create the object $\valv$ at node $n$,
$edit(n,\valv)$---modify the object at $n$ to become $\valv$,
and $remove(n)$---delete the object at node $n$; 
where $n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
edits that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors when combined with other commands are quite different.
Indeed, Bill Zissimopoulos' work 
demonstrated\footnote{Bill Zissimopoulos, personal communication, August 2015.}
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
output \emph{value} of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

Notice, however, that the commands listed above also encode some information about 
their input, the state of the filesystem
before the command is applied. In particular, $create(n,\valv)$ requires that there are no files
or directories at $n$, while $edit(n,\valv)$ and $remove(n)$ require the opposite.
This creates an arbitrary asymmetry where
there is now more information encoded into commands about their output than about their input.
As, based on the above, we expect that encoding more information in the commands
results in a model with greater predictive powers,
and in order to resolve this asymmetry, 
we propose a set of commands that encode
the type of the original state of $\FS(n)$ as well.
(Some real-life filesystem commands like $rmdir$ do this already.)

We therefore have four pieces of information related to each command:
its input type, its output type, the node to which it is applied, and finally
the value that it stores at the node.
We will denote commands using quadruples that list these,
and use $\cchard$, $\ccharf$ or $\ccharb$ to represent the different
input and output types. For example, $\cbdaa{n}{\vald}$ 
will represent $mkdir(n,\vald)$.
Alternatively, we use a triple of the two types and a node if 
there is only a single suitable output value for the given type (e.g. $\empt$).
That is, $\cdba{n}$ will represent $rmdir(n)$.

@@

The exact behavior of the commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxynv$ where
$X$ and $Y$ are types in $\typeset$,
$n\in\setn$ and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
\[ \cxynv\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

@@

% Simplification
% --------------

### 2015-12-22 2015/def-filesystem.tex
More improvements and references

The nodes form a disjoint union of rooted directed trees.
\begin{mydef}
The function $\parent(n)$ returns the parent node of $n$, or
returns $\topnode$ if $n$ is the root of a tree.
\end{mydef}

@@

Another simplification we will assume in this paper is regarding directories.
As Bill Zissimopoulos pointed 
out,\footnote{Bill Zissimopoulos, personal communication, August 2015.} we often do not want to consider metadata stored in
directories during synchronization. In other words, the contents of directories are all equal,
which can be modelled by assuming that $|\setd|=1$.

>>>>

The nodes form a disjoint union of rooted directed trees.
Tao et al. in \cite{TSR} describe a similar filesystem model, although
they also model inodes and restrict the filesystem to a just single tree.
\begin{mydef}
The function $\parent(n)$ returns the parent node of $n$, or
returns $\topnode$ if $n$ is the root of a tree.
\end{mydef}

@@

% Simplification
% --------------

A simplification we will assume in this paper is regarding directories.
As Bill Zissimopoulos pointed 
out,\footnote{Bill Zissimopoulos, personal communication, August 2015.} 
we often do not want to consider metadata stored in
directories (e.g. permission settings) during synchronization,
as these are not generally understood by users,
and, if needed, conflict resolution on these settings can be easily automated.
In other words, the contents of directories are all equal,
which can be modelled by assuming that $|\setd|=1$.

### 2015-12-22 2015/ordering.tex
Minor changes
Minor improvements
More improvements and references

\section{Ordering commands}\label{ordering}

We often encounter the case where we only have a set of commands without a specified order.
As we have seen above, this can occur after the first stage of update detection,
but, more importantly,
it is actually the task of the reconciler to determine whether there is an order,
and if yes, what order,
in which updates from multiple replicas can be applied to a filesystem.

@@

\begin{mylem}\label{connected_changes}
Given a set of commands that
does not contain assertion commands,
and which can be applied to a filesystem in some order without breaking it,
and which contains commands on $n$ and $\nn$ where $\nn\descendant n$
but $\nn\neq\parent(n)$, then the set must also contain a commands
on each node between $n$ and $\nn$.
\end{mylem}
\begin{proof}
We prove that under the given conditions, the set must contain a command on $\parent(n)$.
Then, by repeating this statement, we how that the set must contain commands on every
ancestor of $n$ up to $\nn$.

@@

This is not a problem in the case of components containing a single command only.
(Note that $\cff$ can only appear in a component of one.)

>>>>

\subsection{Ordering commands}\label{ordering}

We often encounter the case where we only have a set of commands without a specified order.
As we have seen above, this can occur after the first stage of update detection,
but, more importantly,
it is actually the task of the reconciler to determine whether there is an order,
and if yes, what order,
in which updates from different replicas can be applied to a filesystem.

@@

\begin{mylem}\label{connected_changes}
Given a set of commands that
does not contain assertion commands,
and which can be applied to a filesystem in some order without breaking it,
and which contains commands on $n$ and $\nn$ where $\nn\descendant n$
but $\nn\neq\parent(n)$, then the set must also contain a command
on each node between $n$ and $\nn$.
\end{mylem}
\begin{proof}
We prove that under the given conditions, the set must contain a command on $\parent(n)$.
Then, by repapplying this result, we how that the set must contain commands on every
ancestor of $n$ up to $\nn$.

@@

This is not a problem in the case of components containing a single command only.
(Note that a file-to-file command ($\cff$) can only appear in a component of one.)

### 2015-12-22 2015/paper.tex
More improvements and references

\section{Comparisons}

We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (see \cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

The former observation is supported by Berzan and Ramsey, who in \cite{CBNR} 
describe between different reconciliation policies.
The liberal (maximal) policy propagates all updates to all replicas where
the command does not break the filesystem, while a conservative policy
refrains from updating any node that is below a node with conflicting commands.
They show that the reconciliation algorithm in \cite{NREC} implements
an intermediate policy as one can easily construct two- or three-replica scenarios
where an update could clearly be propagated, but it is excluded.
To rephrase the example in \cite{CBNR}, consider the following two
update sequences that have been applied to replicas $\FS_A$ and $\FS_B$:
\begin{align*}
A&=\cbdaa{\nn}{\valdi}\cc\cbfaa{n}{\valf} \\
B&=\cbdaa{\nn}{\valdii}
\end{align*}
where $\parent(n)=\nn$ (\cite{NREC} did not require all directories to have the same value).
Clearly $\cbfaa{n}{\valf}$ could be applied to $\FS_B$, but it is not as
it must be preceded by creating the directory, which conflicts with the same operation in $B$
due to the different value.
(Introducing a third replica which has not changed at all further complicates the picture.)
The current algorithm no longer needs to specify that there can be no conflicts
on preceding commands, and, to use the above terminology, implements a fully liberal policy.

% TODO Conclusions

@@

>>>>

\input{comparisons}

@@

\section{Conclusions}

% TODO
Conclusions go here
